A significance test in forward stepwise regression

### 2014-05-04 paper/paper.tex
Finished comparison of stopping rules

\jonathan{Is $r_0=0$ or $y$? Should be defined in the algorithm. I also put a $y$ in below which I'm not sure
is correct or not. }
\josh{Yes, I forgot to specify $r_0 = y$, hence the $y$ is not required (I removed it)}

@@

\begin{figure}
\begin{center}
\subfigure[Categorical design]{
\label{fig:fscat}
\includegraphics[width=0.5\textwidth]{../figs/tpr_default_categorical_nsim100_n100_p400_g200_sizes2-2_k24_lower2_upper4.pdf}}
\hspace{-15pt}
\subfigure[Correlated gaussian design]{
\label{fig:fsgausscorr}
\includegraphics[width=0.5\textwidth]{../figs/tpr_default_gaussian_nsim100_n100_p200_g200_k24_lower1pt2_upper2_corr0pt1.pdf}}
\caption{\small \it The left panel shows the results of the simulation
  using a categorical design consisting of 200 binary factors.
  The right panel shows the results with design matrices of Gaussian
  entries with equi-correlation 0.1 between all columns. Plotted in
  green are the TPP and average model sizes for models chosen by BIC.}
\label{fig:fwdstepsim}
\end{center}
\end{figure}

For various sparsity levels $k$, Algorithm~\ref{algo:fs} was run on
various data sets with \textit{steps} equal to $k$.
In the resulting active set $A_k$, the number of true variables divided by $k$ is the True Positive Proportion (TPP):
\[
\text{TPP}_k = \frac{\# \left\{g: g \in A_k, \beta_g \neq 0\right\}}{k}.
\]
Since we know $k$ in the simulations we can use $\text{TPP}_k$ as a measure of how well forward stepwise is performing. When it is close to 1 we are recovering most of the true variables before including false ones.
Nonzero coefficients had magnitudes
in a range of multiples of $\gamma := \sqrt{2\log(G)/n}$, e.g. in $[1.4
\gamma, 1.6 \gamma]$. Results are shown in
Figure~\ref{fig:fwdstepsim}. After computing the
coherence of these design matrices, some calculations show the required
sparsity level to guarantee exact recovery in these situations is
about 2 or smaller, and the required nonzero coefficient magnitude is
likely in the range of 10 to 100 times $\gamma$. These simulations are far
from the stringent conditions required by theory to guarantee perfect
recovery, but the performance, while not perfect, may still be good
enough for some purposes. Finally, to serve as a comparison with existing
commonly used model selection procedures, in Figure~\ref{fig:fwdstepsim}
we show in green the TPP and average model sizes for models chosen by
the \textit{step} function in \textit{R} using the BIC criterion.

@@

\todo Compare to AIC and other classical stopping rules? \\

@@

\jonathan{make the tick marks denoting the number of true groups larger -- can you not just shade the right hand part gray ? easy to do in python...}
\josh{Done, updated plots now have this form. Will soon finish updating all plots}
Finally, most of the plots below show the following information. The
horizontal axis is the step index for forward stepwise. The dashed
line shows the proportion of iterations where a truly nonzero variable
was added to the model at that step. Solid verticle lines show the
middle 50\% of p-values calculated at that step, with red calculated
on a null model with no nonzero groups, blue calculated on the non-null
model, and green calculated on the non-null model from the usual
$\chi^2$ significance test.
The points in these lines are the average p-value at that step, and
the triangles above and below show the 95\% and 5\% quantiles.

@@

\input{../tables/default_gaussian_nsim500_n100_p50_g50_k5_lower1pt5_upper2.tex}

\input{../tables/default_gaussian_nsim500_n100_p50_g50_k5_lower1pt5_upper2_estimation.tex}

@@

\input{../tables/default_gaussian_nsim500_n100_p200_g200_k5_lower1pt5_upper2.tex}

@@

\begin{figure}
\begin{center}
\subfigure[Forward stepwise with splines]{
\label{fig:spline:fwd}
\includegraphics[width=0.5\textwidth]{../figs/tpr_gamsel_uniform_nsim50_n100_p200_g200_k21_k01_lower5_upper10.pdf}}
\hspace{-15pt}
\subfigure[P-value with splines]{
\label{fig:spline:pval}
\includegraphics[width=0.5\textwidth]{../figs/gamsel_uniform_nsim100_n200_p50_g50_k6_k02_lower10_upper15.pdf}}
\caption{\small \it The left panel shows the power of forward stepwise
for various sparsity levels. The right panel shows our p-value marginally
over each step, with red indicating the global null case and blue
indicating a signal with $k=6$ nonzero groups.}
\label{fig:spline}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\subfigure[Forward stepwise on glinternet]{
\label{fig:glint:fwd}
\includegraphics[width=0.5\textwidth]{../figs/tpr_glint_categorical_nsim50_n100_p72_g36_sizes2-2_k15_k01_lower3_upper5.pdf}}
\hspace{-15pt}
\subfigure[P-value with glinternet]{
\label{fig:glint:pval}
\includegraphics[width=0.5\textwidth]{../figs/glint_gaussian_nsim100_n100_p30_g30_k6_k02_lower4_upper8.pdf}}
\caption{\small \it The left panel shows the power of forward stepwise
for various sparsity levels. The right panel shows our p-value marginally
over each step, with red indicating the global null case and blue
indicating a signal with $k=6$ nonzero groups.}
\label{fig:glint}
\end{center}
\end{figure}

>>>>

%\jonathan{Is $r_0=0$ or $y$? Should be defined in the algorithm. I also put a $y$ in below which I'm not sure
%is correct or not. }
%\josh{Yes, I forgot to specify $r_0 = y$, hence the $y$ is not required (I removed it)}

@@

\begin{figure}
\begin{center}
\subfigure[Categorical design]{
\label{fig:fscat}
\includegraphics[width=0.5\textwidth]{../figs/tpr_default_categorical_nsim100_n100_p400_g200_sizes2-2_k24_lower2_upper4.pdf}}
\hspace{-15pt}
\subfigure[Correlated gaussian design]{
\label{fig:fsgausscorr}
\includegraphics[width=0.5\textwidth]{../figs/tpr_default_gaussian_nsim100_n100_p200_g200_k24_lower1pt2_upper2_corr0pt1.pdf}}
\caption{\small \it The left panel shows the results of the simulation
  using a categorical design consisting of 200 binary factors.
  The right panel shows the results with design matrices of Gaussian
  entries with equi-correlation 0.1 between all columns. Plotted in
  green are the TPP and median model sizes (top) for models chosen by BIC,
  and in red the same are shown (bottom) for RIC.}
\label{fig:fwdstepsim}
\end{center}
\end{figure}

For various sparsity levels $k$, Algorithm~\ref{algo:fs} was run on
various data sets with \textit{steps} equal to $k$.
In the resulting active set $A_k$, the number of true variables divided by $k$ is the True Positive Proportion (TPP):
\[
\text{TPP}_k = \frac{\# \left\{g: g \in A_k, \beta_g \neq 0\right\}}{k}.
\]
Since we know $k$ in the simulations we can use $\text{TPP}_k$ as a measure of how well forward stepwise is performing. When it is close to 1 we are recovering most of the true variables before including false ones.
Nonzero coefficients had magnitudes
in a range of multiples of $\gamma := \sqrt{2\log(G)/n}$, e.g. in $[1.4
\gamma, 1.6 \gamma]$. Results are shown in
Figure~\ref{fig:fwdstepsim}. After computing the
coherence of these design matrices, some calculations show the required
sparsity level to guarantee exact recovery in these situations is
about 2 or smaller, and the required nonzero coefficient magnitude is
likely in the range of 10 to 100 times $\gamma$. These simulations are far
from the stringent conditions required by theory to guarantee perfect
recovery, but the performance, while not perfect, may still be good
enough for some purposes. Finally, to serve as a comparison with existing
commonly used model selection procedures, in Figure~\ref{fig:fwdstepsim}
we show in green the TPP and median model sizes for models chosen by
the \textit{step} function in \textit{R} using the BIC criterion, and
the same for RIC are plotted in red.

@@

@@

Finally, most of the plots below show the following information. The
horizontal axis is the step index for forward stepwise. The dashed
line shows the proportion of iterations where a truly nonzero variable
was added to the model at that step. Solid verticle lines show the
middle 50\% of p-values calculated at that step, with red calculated
on a null model with no nonzero groups, blue calculated on the non-null
model, and green calculated on the non-null model from the usual
$\chi^2$ significance test.
The points in these lines are the average p-value at that step, and
the triangles above and below show the 95\% and 5\% quantiles.

@@

@@

%\input{../tables/default_gaussian_nsim500_n100_p200_g200_k5_lower1pt5_upper2.tex}

\input{../tables/stop_comparison_easy.tex}

\input{../tables/stop_comparison_hard.tex}

@@

\begin{figure}
\begin{center}
\subfigure[Forward stepwise with splines]{
\label{fig:spline:fwd}
\includegraphics[width=0.5\textwidth]{../figs/tpr_gamsel_uniform_nsim100_n100_p200_g200_k18_k01_lower5_upper10.pdf}}
\hspace{-15pt}
\subfigure[P-value with splines]{
\label{fig:spline:pval}
\includegraphics[width=0.5\textwidth]{../figs/gamsel_uniform_nsim100_n200_p50_g50_k6_k02_lower10_upper15.pdf}}
\caption{\small \it The left panel shows the power of forward stepwise
for various sparsity levels. The right panel shows our p-value marginally
over each step, with red indicating the global null case and blue
indicating a signal with $k=6$ nonzero groups.}
\label{fig:spline}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\subfigure[Forward stepwise on glinternet]{
\label{fig:glint:fwd}
\includegraphics[width=0.5\textwidth]{../figs/tpr_glint_categorical_nsim100_n100_p72_g36_sizes2-2_k15_k01_lower5_upper10.pdf}}
\hspace{-15pt}
\subfigure[P-value with glinternet]{
\label{fig:glint:pval}
\includegraphics[width=0.5\textwidth]{../figs/glint_gaussian_nsim100_n100_p30_g30_k6_k02_lower4_upper8.pdf}}
\caption{\small \it The left panel shows the power of forward stepwise
for various sparsity levels. The right panel shows our p-value marginally
over each step, with red indicating the global null case and blue
indicating a signal with $k=6$ nonzero groups.}
\label{fig:glint}
\end{center}
\end{figure}
